\chapter* {Project Description}

\begin {quote}
    \emph {The following text was written as an introduction to Propellerhead
           and the master-thesis project by Erik Agsjö and Gustaf Taxén.}
\end {quote}

Formed in 1994, Propellerhead (\url {http://www.propellerheads.se/}) is a
privately owned company based in Stockholm, Sweden. Renowned for its
musician-centric approach, Propellerhead has created some of the world's most
innovative music software applications, hardware interfaces and technology
standards.

Musicians, producers and the media have praised \emph {Reason}, \emph {ReCycle}
and \emph {ReBirth} applications for being inspiring, great sounding and of
superior quality. Technologies such as \emph {ReWire} and the \emph {REX} file
format are de-facto industry standards, implemented in all major music software.
Today, Propellerhead’s products are used all over the world by hundreds of
thousands of professionals and enthusiasts for all kinds of music making.

In the summer of 2012, Propellerhead launched one of the most significant
updates to the Reason software: \textbf {Rack Extensions}. Rack Extensions are
third-party plug-ins for Reason that extends the capabilities of the Reason host
in various ways. Examples include (virtual) instruments, sound effects
processors, and audio analysis utilities. All Rack Extensions are sold
exclusively in the Propellerhead on-line shop,
\url {http://shop.propellerheads.se/}.

Propellerhead focuses very strongly on quality and robustness and the Rack
Extension concept is no exception. Once a Rack Extension has been approved by
Propellerhead, it will be supported in Reason indefinitely, and songs created
with an old version of a Rack Extension will continue to work even if the Rack
Extension is updated by its developer. Each Rack Extension runs in a protected
sandbox environment inside Reason, which ensures that the user does not lose any
data if the Rack Extension should crash. Rack Extensions are also highly
portable and the format assumes very little about the operating environment in
which the Reason host is running (Reason is currently available for Mac OSX and
Windows, in both 32- and native 64-bit versions).

Real-time audio processing is computationally expensive, so Rack Extensions are
written in C++ for efficiency reasons. However, third-party developers do not
submit C++ source code to Propellerhead. Instead, we use the portable byte code
format of the LLVM compiler. This ensures that the implementation details of the
third party code are protected, while allowing Propellerhead to maintain the
submitted code indefinitely, without having to keep in contact with the original
developer.

In summary, Rack Extension plugins are \emph {portable},
\emph {computationally efficient}, and \emph {robust}. The Reason host already
provides a very strong environment with respect to these requirements, but there
are a number of areas that could be improved. This Master’s Thesis focuses on
some of these areas.
